---
title: "Exponential Distribution and the Central Limit Theorem"
author: "James Whitehead"
date: "29 December 2016"
output: pdf_document
---

##Synopsis

In this report, the exponential distribution is investigated and compared to the Central Limit Theorem (CLT). A large number of exponential distributions are simulated, and the means are shown to follow an approximate normal distribution.

##Simulations

Sets of 40 exponential distributions with lambda = 0.2 as the rate constant are simulated 1000 times and stored as a matrix.

```{r}
set.seed(1)
expDist <- matrix(rexp(40*1000,0.2), 1000, 40)
expDistMeans = apply(expDist, 1, mean)
mean(expDistMeans)
hist(expDistMeans, main="Means of exponential distributions", xlab="Mean")
abline(v=5, lwd=5)
```

The figure above shows the distribution of means of the exponential distributions. The thick vertical line shows the theoretical mean of 5 (1/lambda), which is close to the sample mean calculated above of 4.99. The theoretical variance of the distribution is (1/lambda)^2/n = 0.025. For the simulations:

```{r}
expDistVar = apply(expDist, 1, var)
sum(expDistVar)/1000^2
```


Finally, to show that the distribution of means is approximately normal, the distribution generated above is compared with an exponential distribution of 1000 simulated variables:

```{r}
par(mfcol=c(1,2), mar=c(5,4,4,4))
hist(expDistMeans, main="Means of exponential distributions", xlab="Mean", prob=TRUE, 
     ylim=0:1)
curve(dnorm(x, 5, 0.62), add=TRUE, col="red", lwd=2)
hist(rexp(1000,0.2), main="Exponential distribution", xlab="exp")
```

This shows that the distribution of means is much closer to Gaussian than the exponential distribution, as illustrated by the overlaid red curve.